Applying Gaussian Process Regression to Predict Changes in Stock Price

### 2014-05-10 paper/paper.tex
latest status

\subsection{Pre-processing}

@@

Then, the covariance functions are:
\begin{align}
	k(\bfx_p, \bfx_q) &= \sigma_f^2 \exp{- \frac{1}{2} u^2} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{Squared Exponential}   \\
	k(\bfx_p, \bfx_q) &= \sigma_f^2 \exp{- u} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{Exponential}   \\
	k(\bfx_p, \bfx_q) &= \sigma_f^2 \left( 1 + \sqrt{3} u \right) \exp{- \sqrt{3}  u} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{Matern ($\nu = \frac{3}{2}$) }   \\
	k(\bfx_p, \bfx_q) &= \sigma_f^2 \left( 1 + \sqrt{5} u \right) + 5 u^2) \exp{- \sqrt{5} u} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{Matern ($\nu = \frac{5}{2}$) }
\end{align}

@@

\section{Conclusion}

The results from our experiments indicate that GP models could learn the correlations hidden among financial data to a very high degree when compared to other traditional ML algorithms like linear regression and PCA regression. But the biggest drawback of GP models as implemented in our application is the computation time that is required for the GP modeling. Regression based on GP models involve a huge amount of matrix inversions, the complexity of which is to the order of $\mathcal{O}(n^3)$. These computational needs, restricted the amount of training data that could be used, to a maximum of only 20000 training samples, even though we had access to around 2GB of stock price data that we would have liked to use during training. Several authors have suggested sparse approximation approaches for calculating the covariance matrix, which we believe is a promising direction for future work.

>>>>

\subsection{Pre-processing}
\label{sec:strategy}

@@

Then, the covariance functions are:
\begin{align}
	k(\bfx_p, \bfx_q) &= \sigma_f^2 \exp{- \frac{1}{2} u^2} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{(Squared Exponential)}   \\
	k(\bfx_p, \bfx_q) &= \sigma_f^2 \exp{- u} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{(Ornstein–Uhlenbeck)}   \\
	k(\bfx_p, \bfx_q) &= \sigma_f^2 \left( 1 + \sqrt{3} u \right) \exp{- \sqrt{3}  u} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{(Matern - $\nu = \frac{3}{2}$)}   \\
	k(\bfx_p, \bfx_q) &= \sigma_f^2 \left( 1 + \sqrt{5} u \right) + 5 u^2) \exp{- \sqrt{5} u} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{(Matern  - $\nu = \frac{5}{2}$)}
\end{align}

@@

\section{Experiments}

\subsection{Covariance Type}
For this experiment, we use $\mathcal{T} = 20000$, with Strategy 3, and try to predict 500 ms into the future.
We found that squared exponential and the Matern function with $\nu = \frac{5}{2}$ predicted nearly a straight line, with very small variations.
The Matern functio wiht $\nu = \frac{3}{2}$ and Ornstein–Uhlenbeck functions' predictions are shaped more like the actual data, with sharp peaks and valleys.

The Ornstein-Uhlenbeck model seems to fit the data the best, so we use it for all future experiments.

% Zoom plot in - 50-100??

\subsection{Window size}
Next, we experimented with varying window size $\mathcal{T}$, while using Strategy 3 and predicting 500 ms into the future.
We found that the larger windows predicted the dominant trend much better, and had fewer peaks and valleys.
However, the peaks and valleys that the larger windows predicted were, on average, more likely to be correct.

\subsection{Strategy Type}
We tried several different strategies (described in \ref{sec:strategy}).
\begin{table}[h]
	\label{t:miss}
	\centering
	\caption{Earnings from different strategies}
	\begin{tabular}{| l | l |}
	\hline
	Strategy & Earnings \\
	\hline
	1 & 42.88 \\
	2 & 28.39 \\
	3 & 3.86 \\
	\hline
	\end{tabular}
\end{table}

\subsection{Future time}
In experiments up to this point, we used data that had been processed so that the y values are the change in stock price 500 ms in the future.
Next, we re-processed the data to have y values 800 ms or 1000 ms in the future.
We found that we got roughly the same results for each experiment, suggesting that we are equally capable of predicting on each of these time scales into the future.

If we had more time, we would run additional experiments further into the future, to see how far ahead we can predict before our predictions get notably worse.

\subsection{Comparisons to other models}
The dataset that we used came with linear regression and PCA regression models, trained on a much larger version of the same dataset.
Although the experiments are not trained and tested in the same way, we give a broad comparison to their results, to show that we performed similarly well.

\section{Conclusion}
The results from our experiments indicate that GP models could learn the correlations hidden among financial data to a very high degree when compared to other traditional ML algorithms like linear regression and PCA regression. But the biggest drawback of GP models as implemented in our application is the computation time that is required for the GP modeling. Regression based on GP models involve a huge amount of matrix inversions, the complexity of which is to the order of $\mathcal{O}(n^3)$. These computational needs, restricted the amount of training data that could be used, to a maximum of only 20000 training samples, even though we had access to around 2GB of stock price data that we would have liked to use during training. Several authors have suggested sparse approximation approaches for calculating the covariance matrix, which we believe is a promising direction for future work.
